---
title: "scWGCNA workflow"
output:
  html_document:
    fig_height: 7
    fig_width: 8
    df_print: paged
  pdf_document:
    fig_height: 7
    fig_width: 8
    df_print: paged
editor_options: 
  chunk_output_type: inline
params:
  data: No_data
  sc.data: No_dataproject
  project.name: No_name
  gene.names: No_names
  cells: No_cells
  features: No_features
  reduction: No_reduction
  dir: No_dir
  sp: No_sp
  is.pseudocell: No_info
  GO: No_GO
  min.cell: 10
---

Date: `r format(Sys.Date(), "%d.%m.%y")`

We use WGCNA to run an iterative analysis on a data set. 
If no features are prided, HVGs are calculated in the script in the script, using Seurat, the mvp method a threshold of 0.25 for the dispersion.      

## Pre-analysis

We need to first set up our working environment.

```{r Set up, message=FALSE, warning=FALSE, results='hide'}

my.date = format(Sys.Date(), "%d.%m.%y")

# The following setting is important, do not omit
backup.options = options()
options(stringsAsFactors = FALSE)

```

Parse all the parameters and data we need to be able to run the analysis.
We need the data in seurat format. This must contain at least 20 samples (no problems with sc data), according to the [documentation of the package itself](https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/faq.html).

```{r From above, message=FALSE, warning=FALSE, results='hide'}

# placeholder for the variables and options we will get from an upper level pipeline

project_name = params$project.name

p.Wdata = params$sc.data

gnames = params$gene.names
gnames[,2]= as.character(gnames[,2])
rownames(gnames) = gnames[,1]

Wdata = params$data

# The subset in form of cell identities, F if using the whole sample
my.subset = params$cells
# Variable genes, or genes to use, F if the genes will be calculated in the script
my.vargenes = params$features

my.dr = params$reduction

dir.create(paste0(params$dir,"./WGCNA_results"), showWarnings = F)

my.sp = params$sp

is.pseudocell = params$is.pseudocell

my.min.cell = params$min.cell

```

Now we calculate the genes we will use.

```{r pre-process the data}

# If we need to subset the data
if (my.subset) {
  p.Wdata = subset(p.Wdata, idents=my.subset)
}

# If no variable genes are provided
nonex = which(apply(p.Wdata@assays$RNA@counts, 1, function(x) length(which(x >0))) < my.min.cell)

if (my.vargenes == F) {
  # First get rid of non-expressed genes
  p.Wdata = subset(p.Wdata, features = rownames(p.Wdata@assays$RNA)[-nonex])
  
  # Find the variable genes
  p.Wdata=Seurat::FindVariableFeatures(p.Wdata, dispersion.cutoff = c(0.25,Inf), mean.cutoff = c(0,Inf), selection.method = "mvp", assay = "RNA")
  Seurat::VariableFeaturePlot(p.Wdata, assay = "RNA")
  Expr = Seurat::VariableFeatures(object = p.Wdata, assay = "RNA")
  
} else { Expr = my.vargenes }

# if (length(my.ortho) > 1) {
#   Expr = Expr[which(Expr %in% my.ortho[,1])]
#   Expr = Expr[ which(Expr %in% rownames(p.Wdata@assays$RNA)[-nonex]) ]
# }
 
if (is.pseudocell) {
  
  datExpr=Wdata@assays$RNA@counts[Expr,]
  datExpr = datExpr[which(rowSums(as.matrix(datExpr))>0),]
  Expr = rownames(datExpr)
  
} else{datExpr=Wdata@assays$RNA@data[Expr,]}

# Check the length
print(paste0("We have ", length(Expr), " genes in the variable genes object"))

# Check the size and transform
dim(datExpr)
datExpr = t(as.matrix(datExpr))

```

Now we need to calculate the soft threshold power. First it calculates the similarity and then transforms this similarity to a weighted network. The scale-free topology is calculated for each of the powers.

We choose the smallest power for which the scale-free topology fit index reaches 0.90. 
If none of the powers reaches 0.90, we take the one with the maximum, as long as we have a number above 0.75. If none of them reaches at least 0.75 we need to check our dataset.

```{r soft-thresholding power, echo=FALSE, message=FALSE}

# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to=30, by=2))

# Call the network topology analysis function
sft = WGCNA::pickSoftThreshold(datExpr, powerVector = powers, verbose = 0, networkType = "signed", corFnc = "bicor")

# Plot of the scale-free topology fit index as a function of the soft-thresholding power
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
     main = paste("Scale independence"))
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
     labels=powers,col="red")
which((-sign(sft$fitIndices[,3])*sft$fitIndices[,2]) > 0.9)
# this line corresponds to using an R^2 cut-off of h
abline(h=0.85,col="red")

# These are the scale-free topology indexes
indexes = (-sign(sft$fitIndices[,3])*sft$fitIndices[,2])

# If we don't have aany index above 0.75, we stop the script
if ( !any( indexes > 0.75 ) ) {
  print("The scale-free topology index didn't reach 0.75 with any of the chosen powers, please consider changing the set of genes or cells provided")
  # quit(save = "no", 1, F)
}

if ( !any( indexes > 0.4 ) ) {
  print("The scale-free topology index didn't reach 0.4 with any of the chosen powers, the script will not continue, since it might fail")
  
  knitr::knit_exit()

}

# Take the smalles power that gives us an index over 0.9, or the highest index if we don't reach 0.9
if ( any( indexes > 0.9 ) ) {
    my.power = sft$fitIndices$Power[min(which(indexes > 0.9))]
} else { my.power = sft$fitIndices$Power[which.max(indexes)] }

print(paste0("Or power is ", my.power))


```

## WGCNA analysis

Now, the following piece of code will run WGCNA iteratively, to end up with out final modules.
The iterations follow these steps:

- Tree cutting for module calculation
  - Calculate an adjacency matrx from the data, then turn it into topological overlap and then into a distance matrix
  - Calculate the tree based on the topological overlap distance
  - Calculate the automatic height to cut out the 0.05 quantile
  - Generate a matrix where we calculate the amount of modules, size of de modules and wheter if we have a grey module based on:
    - Different minimum module sizes, arbitrarely set to 7:30 
    - Different cut-heiughts going 0.0005 up and down from the automatic height in steps of 0.0001
  - Check if any combination of the parameters will get rid of the grey module
    - If we only get grey modules, subset the matrix for the height at which the grey module is the smallest
    - If we have a combination without grey module AND we have at least the same number of modules as in the beginning, we subset for whichever those heights are
  - Take wichever min module sizes gives us at least the same amount of clusters as in the beginning, if none, then the highest
  - Chose the max of the reminding min module sizes.
- Calculate the actual modules
- Calculate the eigengenes
- Calculate the module membership per gene
- Calculate the p.value of the membership to a given module
- Get rid of the genes in the grey module
- Delete the genes that are not significantely associated with their module
- Save the remaining geneset
- Print how many genes were deleted due to significance
- Unless 0 genes were deleted in the last step, update the expression matrices and beginn again.

ONLY THE FIRST TWO ITERATIONS ARE DIFFERENT.

FIRST:

- During the tree cutting
  - Set the minimum module size to arbitrary 15
  - Subset for the heights that get rid of at least 50% of the genes with that min module size
  - Choose the height that gives us the most clusters

SECOND:

- Set the resulting number of modules as the ground number of modules

```{r WGCNA}
# Running WGCNA iteratively

my.Clnumber = 20
change = 0
genesets=list()
nonsig = 1

while(nonsig != 0) {
  
  my.adjacency =WGCNA::adjacency(datExpr,type = "signed", power = my.power, corFnc = "bicor")
  
  # Turn adjacency into topological overlap (high overlap if they share the same "neighborhood")
  TOM=WGCNA::TOMsimilarityFromExpr(datExpr,networkType = "signed", TOMType = "signed", power = my.power, corType = "bicor")
  
  #Put the names in the tree
  colnames(TOM) <- gnames[colnames(datExpr),2]
  
  rownames(TOM) <- gnames[colnames(datExpr),2]
  
  #Make it a distance
  dissTOM = 1-TOM
  
  # Call the hierarchical clustering function
  geneTree = hclust(as.dist(dissTOM), method = "average")
  
  
  # Here I calculate the cutting height. Using the same formula and approach that the WGCNA package uses for the automatic function
  nMerge = length(geneTree$height) # The whole height of the tree
  refQuantile = 0.05 # What's the quantile that we want to exclude
  refMerge = round(nMerge * refQuantile) 
  refHeight = geneTree$height[refMerge]
  cutheight = signif(0.99 * (max(geneTree$height) - refHeight) + refHeight,4)
  
  # We construct THE TABLE that will help us make decisions
  # Min cluster sizes, from 7 to 30
  x=seq(7,30,1)
  # The height, up and down from the calculated height. We expect some "No module detected"
  y=seq(cutheight-0.0005,cutheight + 0.0005,0.0001)
  
  # The actual dataframe
  w=data.frame()
  # Populate, with i=min cluster size, j=cutting height, z=total number of clusters, z.1.=what's the first cluster? 0 is grey 1 is something else,
  # z.1.'=what's the size of the first cluster?
  for (i in x) {
    for (j in y) {
      sink("aux")
      z=table(dynamicTreeCut::cutreeDynamic(dendro = geneTree,  method="tree", minClusterSize = i, deepSplit = T, cutHeight = j, verbose = 0))
      sink(NULL)
      v=data.frame(i,j,dim(z),names(z[1]),unname(z[1]))
      w=rbind(w,v)
    }
  }
  
  # The height is then the one where we have the least number of genes in the first cluster
  my.height = w$j[which(w$unname.z.1..==min(w$unname.z.1..))]
  
  # Since different heights can give us the minimum grey size, we chose the computed height, if present, or the highest one.
  if (cutheight %in% my.height) {
    my.Clsize = w[which(w$j == cutheight),]
  } else { my.Clsize = w[which(w$j == max(my.height)),] }
  
  
  # This is to know, if we're looking for a minimum of cluster numbers
  # If we still have a lot of genes, we don't want to limit the number of clusters
  # if ( ((dim(datExpr)[2]) / length(Expr)) > 0.6 ) {change = 0}
  
  # If this is the first iteration after 0.6 of the genes are gone, we assign the number of clusters (and an extra for the grey in the case)
  if (change == 2) {
    my.Clnumber = length(table(dynamicColors)) + 1
  }
  # Count another iteration
  change = change + 1
  
  # If we don't have a gray cluster anymore, then we subset for those rows, and set a new height. ONLY if we get the same amount of clusters!
  if (any(w$names.z.1.. == 1)) { #any combination gives us no grey
    
    my.Clsize = w[which(w$names.z.1.. == 1),] # Take all combinations that gives us no grey
    
    if (any(my.Clsize$dim.z. >= (my.Clnumber -1) )) { # If there is any giving us the determined amount or more
      my.Clsize = my.Clsize[which(my.Clsize$dim.z. >= (my.Clnumber - 1) ),,drop=F] # Subset for those
    } else { my.Clsize = my.Clsize[which(my.Clsize$dim.z. == max(my.Clsize$dim.z.)),,drop=F] } # Or for the highest
    
    # Take the ones with the smallest number of clusters
    my.Clsize = my.Clsize[which(my.Clsize$dim.z. == min(my.Clsize$dim.z.)),,drop=F]
    # Take the one with the highest min cluster size
    my.Clsize = my.Clsize[which(my.Clsize$i == max(my.Clsize$i)),,drop=F]
    
    if (cutheight %in% my.Clsize$j) { # if original computed height is in,
      my.height = cutheight # take it
    } else { my.height = max(my.Clsize$j) } # Otherwise, the highest height
    
    my.Clsize = max(my.Clsize$i)
    
  }
  
    # Subset the table again, for those sizes that will gives the same number of clusters or more. IF NONE, use the highest number
  if (!any(w$names.z.1.. == 1)){
    if (any(my.Clsize$dim.z. >= my.Clnumber)) {
      my.Clsize = my.Clsize[which(my.Clsize$dim.z. >= my.Clnumber),,drop=F]
    } else {
      my.Clsize = my.Clsize[which(my.Clsize$dim.z. == max(my.Clsize$dim.z.)),,drop=F]}
    
   # Take the ones with the smallest number of clusters
    my.Clsize = my.Clsize[which(my.Clsize$dim.z. == min(my.Clsize$dim.z.)),,drop=F]
    # Take the one with the highest min cluster size
    my.Clsize = my.Clsize[which(my.Clsize$i == max(my.Clsize$i)),,drop=F]
    
    if (cutheight %in% my.Clsize$j) { # if original computed height is in,
      my.height = cutheight # take it
    } else { my.height = max(my.Clsize$j) } # Otherwise, the highest height
    
    my.Clsize = max(my.Clsize$i)
  }
  
  # If we still have more than 60% of the genes, we just use the min size of 15 regardles
  if ( change < 3 ) {
    my.Clsize = 15
    my.height = cutheight
  }
  
  dynamicMods = dynamicTreeCut::cutreeDynamic(dendro = geneTree,  method="tree", minClusterSize = my.Clsize, deepSplit = T, cutHeight = my.height)
  
  #dynamicMods = cutreeDynamic(dendro = geneTree, distM = dissTOM, deepSplit = 2, minClusterSize = minModuleSize)
  
  table(dynamicMods)
  
  # Convert numeric lables into colors
  dynamicColors = WGCNA::labels2colors(dynamicMods)
  table(dynamicColors)
  
  # Plot the dendrogram and colors underneath
  WGCNA::plotDendroAndColors(geneTree, dynamicColors, "Modules",
                      dendroLabels = NULL,
                      cex.dendroLabels = 0.6,
                      addGuide = TRUE,
                      main = "Gene dendrogram and module colors",
                      guideAll = F)
  
  par(mfrow=c(1,1))
  
  # Calculate eigengenes
  MEList = WGCNA::moduleEigengenes(as.matrix(datExpr), colors = dynamicColors)
  
  MEs = MEList$eigengenes
  
  # Calculate the module membership
  geneModuleMembership = as.data.frame(WGCNA::signedKME(datExpr, MEs))
  MMPvalue = as.data.frame(WGCNA::corPvalueStudent(as.matrix(geneModuleMembership), nrow(datExpr)))
  
  # We're gonna make a list, where we keep the genes that are significantly associated with each module
  x=c()
  xy=list()
  
  # We also need a vector with all the dynamic colors
  dcols = 1:length(levels(as.factor(dynamicColors)))
  
  # Getting rid of the grey module
  grey.genes = length(which(dynamicColors == "grey"))
  if (any(levels(as.factor(dynamicColors)) == "grey")) {
    dcols = dcols[-which(levels(as.factor(dynamicColors)) == "grey")]
  }
  
  # Run the loop to get the genes
  for (i in dcols) {
    modGenes = rownames(MMPvalue)[which(dynamicColors==levels(as.factor(dynamicColors))[i] & MMPvalue[,i]<0.01)]
    x=c(x,modGenes)
    xy[[i]]=modGenes
    #print(paste0(levels(as.factor(dynamicColors))[i]," ",length(modGenes),
    #" of ", length(which(dynamicColors==levels(as.factor(dynamicColors))[i]))))
    #print(gnames[modGenes,2])
  }
  
  # Make a new list, where we keep ALL the gens thar are left from the iteration, that will be used to make the new object. To keep track
  genesets[[length(genesets)+1]] = colnames(datExpr)
  
  # Give me a message saying how many genes are gone this time
  cat( paste0( grey.genes, " genes not assigned to any module.", '\n',
                 length(which(!(colnames(datExpr)%in%x))) - grey.genes, " genes excluded due to significance."))
  # Save this also, cause if it's 0 then we stop the whole thing
  nonsig = length(which(!(colnames(datExpr)%in%x)))
  
  # If it ain't 0, subset the dynamic colors and the expression data
  if (length(which(!(colnames(datExpr)%in%x))) != 0) {
    dynamicColors=dynamicColors[-which(!(colnames(datExpr)%in%x))]
    datExpr=datExpr[,-(which(!(colnames(datExpr)%in%x)))]
  } 

}
```

We calculate here the expression of genes using the single-cell data

```{r raw eigengenes}

p.MEList = MEList

raw.datExpr = p.Wdata@assays$RNA@data[colnames(datExpr),]

raw.datExpr = t(as.matrix(raw.datExpr))

raw.MEList = WGCNA::moduleEigengenes(raw.datExpr, colors = dynamicColors)
  
p.MEList = raw.MEList

```

## Modules of co-expression

For single cells
We can see what are the expression levels of our co-expression modulesin the single cells. Using the dimensionality reduction provided. Default is a tSNE  

```{r Plot multiexpression}
 xx=list()
 yy=c(levels(as.factor(dynamicColors)))
for (i in 1:length(yy)) {
  toplot = data.frame(p.Wdata@reductions[[my.dr]]@cell.embeddings)

  xx[[i]] = ggplot2::ggplot(toplot[order(p.MEList$averageExpr[,i]),],
                   ggplot2::aes_string(x=colnames(toplot)[1], y=colnames(toplot)[2])) +
    ggplot2::geom_point(ggplot2::aes_string(color=p.MEList$averageExpr[order(p.MEList$averageExpr[,i]),i]), size=2) +
    ggplot2::scale_size(range = c(1, 1)) +
    ggplot2::theme_void() +
    ggplot2::theme(legend.position="none") +
    ggplot2::scale_colour_gradientn(colours = c("gray90", "gray90", yy[i], yy[i])) +
    ggplot2::labs(colour=levels(as.factor(dynamicColors))[i])
}
gridExtra::grid.arrange(grobs=xx, ncol=4)
```

Create a list object with all the essential data from the analyses

```{r Save object}

WGCNA_data = list()
WGCNA_data[["datExpr"]] = datExpr
WGCNA_data[["dynamicMods"]] = dynamicMods
WGCNA_data[["MEList"]] = MEList
WGCNA_data[["MEs"]] = MEs
WGCNA_data[["modGenes"]] = xy
WGCNA_data[["modPlots"]] = xx
WGCNA_data[["genesets"]]= genesets
WGCNA_data[["TOM"]]= TOM
WGCNA_data[["adjacency"]]= my.adjacency

saveRDS(WGCNA_data, file = paste0(params$dir,project_name,"_WGCNA_data_",my.date,".rds"))

```

To create the visualizations of the actual co-expression networks, we use this code. It also creates files that can be read into cytoscape

```{r Network plotting, message=F}
# we create a new directory where the network files will be
my.dirname = paste0(params$dir,project_name,"WGCNA_networks_",my.date,"/")

dir.create(my.dirname)

modules=unique(dynamicColors) # The modules

# To generate the files necessary for the network visualization in cytoscape
for (i in 1:length(modules)) {
  mod=is.finite(match(dynamicColors, modules[i])) # Each module
  cyt = WGCNA::exportNetworkToCytoscape(TOM[mod,mod],
                                 edgeFile = paste( paste0(my.dirname,"CytoscapeInput-edges-"),
                                                  paste(modules[i], collapse="-"), ".txt", sep=""), # The name of the edge file
                                 nodeFile = paste( paste0(my.dirname,"CytoscapeInput-nodes-"),
                                                  paste(modules[i], collapse="-"), ".txt", sep=""), # The node file
                                 weighted = TRUE, #Puts the weights
                                 threshold = 0.0, #Threshold for adjacency, 0=all genes
                                 nodeNames = colnames(datExpr)[mod], #The names of the nodes/genes
                                 altNodeNames = gnames[colnames(datExpr)[mod],2], #Other names
                                 nodeAttr = geneModuleMembership[colnames(datExpr)[mod],i]) #Some more info about the nodes
}

# We build a list where we keep the network plots
netplots=list()

lcols = c(rep("black", length(unique(dynamicColors))))
lcols[(apply(col2rgb(levels(as.factor(dynamicColors))), 2,
               function(x) (x[1]*0.299 + x[2]*0.587 + x[3]*0.114)) < 75)] = "white"

mynets = list()

# A loop that makes the network plots in R
for (i in 1:length(xx)) {
  # Load directly edges tables from the file we created
  mynetwork = read.table(paste0(my.dirname,"CytoscapeInput-edges-",levels(as.factor(dynamicColors))[i],".txt"),
                       header = T, stringsAsFactors = F, fill=T)
  # We get rid of all the edges that have a very small weight. The cutoff set to keep at least ONE edge on the nodes
  x = max( c( min(aggregate(mynetwork$weight, by = list(mynetwork$fromNode), max)$x),
              min(aggregate(mynetwork$weight, by = list(mynetwork$toNode), max)$x) ) )
  mynetwork=mynetwork[-which(mynetwork$weight < x),]

  # We rescale the weights so that we have them from 0 to 1
  mynetwork$weight01 = GGally::rescale01(mynetwork[,3])
  # And then multiply them by 2, to give the heavy edges a 2 thickness, we ad 0.2 to not have completelly invisible edges
  mynetwork[,3] = (GGally::rescale01(mynetwork[,3]) * 2) + 0.2
  # Convert this edgelist into a network object
  mynet = network::network(mynetwork, matrix.type="edgelist", ignore.eval=F)

  # Load in the nodes table
  mynodes = read.table(paste0(my.dirname,"CytoscapeInput-nodes-",levels(as.factor(dynamicColors))[i],".txt"),
                     header = T, stringsAsFactors = F, fill = T)
  rownames(mynodes) <- mynodes$nodeName
  # put in the membership value, from the ModuleMembership table, and rescale that from 0 to 1.
  mynodes$membership = GGally::rescale01(geneModuleMembership[rownames(mynodes),i])
  # We multiply by 30 to give a good range of sizes, plus 1 to avoid innexistant nodes
  mynodes$membership= (mynodes$membership*30)+1

  mynet = network::set.vertex.attribute(mynet, "membership", mynodes[network::network.vertex.names(mynet),"membership"])

  mynets[[i]] = mynet

}

```

To check the results module by module, we report some GO terms and individual plots, with module sizes and gene names. First the GO terms analyses

```{r GO analysis, fig.height=10, fig.width=6, message=FALSE}
if (params$GO==T) {
  #Get the ENSEMBL to ENTREZ list from the BioMart

  require(paste0("org.",my.sp,".eg.db"),character.only = T)
  my.ENS2EG = get(paste0("org.",my.sp,".egENSEMBL2EG"), pos = paste0("package:org.",my.sp,".eg.db"))
  
  IDs=as.list(my.ENS2EG)

  #Create two lists, one for the DEGs and one for the GOs
  dewg = list()
  gowg = list()

  # A loop that goes through the DEG files we created earlier
  for(i in 1:length(yy)){ #as many clusters as we have
     dewg[[i]] = xy[[i]] #Read the table
     ex = sapply(dewg[[i]], function(x) AnnotationDbi::exists(x, my.ENS2EG)) #Which genes have an ENTREZ?
    dewg[[i]] = dewg[[i]][ex] #Only those genes
    dewg[[i]] = unlist(IDs[dewg[[i]]]) #The ENTREZ IDs
    gowg[[i]] = limma::goana(dewg[[i]], species = my.sp, universe = unlist(IDs[rownames(Wdata)])) #The GO analysis
   }

  goplots=list()
  goterms=list()

  for (i in 1:length(yy)) {
    toplot=limma::topGO(gowg[[i]], n= 50, ontology = c("BP"))

    #Change to character and then back to factor, to keep the order from TopGO
    toplot$Term = as.character(toplot$Term)
    toplot$Term = factor(toplot$Term, levels = unique(toplot$Term))

    goterms[[i]]=toplot

  }

  #TO make the acutal plot we're having, we combine all the GOs
  g.plot = data.frame()
  topterms = data.frame()

  for (i in 1:length(goterms)) {
    topterms = goterms[[i]][1:5,]
    topterms$cluster = i
    g.plot = rbind(g.plot, topterms)
  }

  g.plot$cluster = as.factor(g.plot$cluster)

  my.terms = nchar(as.character(g.plot$Term))>45
  my.gterms = strtrim(g.plot$Term, 45)
  my.gterms[my.terms] = paste0(my.gterms[my.terms],"...")

  ggplot2::ggplot(g.plot, ggplot2::aes(x=Term, y=-log10(P.DE), fill=cluster)) +
      ggplot2::geom_point(ggplot2::aes(shape = cluster), size = 3) +
      ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, hjust = 1)) +
      ggplot2::labs(x="GO Term", y="-log10 p. value") +
      ggplot2::scale_x_discrete(labels=my.gterms) +
      ggplot2::scale_fill_manual(name = "Module",
                        labels = yy,
                        values = yy) +
      ggplot2::scale_shape_manual(name = "Module",
                         labels = yy,
                         values = rep(c(21,22,23,24,25),20)) +
    ggplot2::coord_flip()
}

```

And here the report of each of the modules

```{r results per module, results="asis", fig.height=9}

# Go trhough the modules
for(i in 1:length(xx)){

  # A table is done, where we put the genes names that are making-up the module, in rows of 10. We fill the las row with empty spaces
  ktable1 = data.frame( matrix( c(gnames[xy[[i]],2],rep(" ", 10-length(gnames[xy[[i]],2])%%10 )), ncol = 10, byrow = T ) )
  colnames(ktable1) = rep(".", 10)

  if (params$GO == T) {
    # Another table, where we put the top 10 GO terms, in two columns
    ktable2 = data.frame( matrix( as.character(goterms[[i]][1:10,1]), ncol = 2, byrow = F ) )
    colnames(ktable2) = rep(".", 2)
    }

  # We make a plot in which we show the mean expression level of the co-expression module

  toplot = data.frame(p.Wdata@reductions[[my.dr]]@cell.embeddings)

  myplot = ggplot2::ggplot(toplot[order(p.MEList$averageExpr[,i]),],
                   ggplot2::aes_string(x=colnames(toplot)[1], y=colnames(toplot)[2])) +
      ggplot2::geom_point(ggplot2::aes_string(color=p.MEList$averageExpr[order(p.MEList$averageExpr[,i]),i])) +
      ggplot2::scale_size(range = c(1, 1)) +
      ggplot2::theme_void() +
      ggplot2::scale_colour_gradientn(colours = c("gray90", "gray90", yy[i], yy[i])) +
      ggplot2::labs(colour=levels(as.factor(dynamicColors))[i])

  mynet=mynets[[i]]

  mynetplot = GGally::ggnet2(mynet,
                         mode = "fruchtermanreingold",
                         layout.par = list(repulse.rad=network::network.size(mynet)^1.1,
                                           area=network::network.size(mynet)^2.3), # Give space in the middle
                         node.size = network::get.vertex.attribute(mynet,'membership'), max_size = 22, #The size of the nodes
                         node.color = levels(as.factor(dynamicColors))[i], # The color of the module
                         edge.size = "weight",
                         edge.color = "black",
                         edge.alpha = network::get.edge.attribute(mynet,'weight01'),
                         ) +
    ggplot2::theme(legend.position="none") +
    ggplot2::geom_label(ggplot2::aes(label=gnames[network::network.vertex.names(mynet),2]),
               fill = levels(as.factor(dynamicColors))[i],
               alpha = 0.5,
               color=lcols[i],
               fontface = "bold")

  # We put our mean expression plot toghether with the network plot
  gridExtra::grid.arrange(grobs=list(myplot,mynetplot), nrow=2)
  # # plot.new()
  # # dev.off()
  cat("\n")
  # A table showing the number, color and size of the module
  print(knitr::kable(data.frame(module=i, color=names(table(dynamicColors))[i], size=unname(table(dynamicColors)[i]))))
  # cat("\n")
  # The other two tables we just made
  print(knitr::kable(ktable1))
  if (params$GO == T) {
   print(knitr::kable(ktable2))
  }
  cat("\n")


}

```

```{r info}

options(backup.options)

sessionInfo()

```